### 2017-05-10 paper2.tex
Add files via upload

\subsection{Evaluation and parameterization}

@@

By virtue of requiring nearly 500 radiative transfer simulations per time step and significant post-processing, the method described in \ref{method2} is computationally expensive. This prompted the development of a parameterization scheme to facilitate more widespread usage of urban T\textsubscript{hem, r} for climatological and comparative sUHI analysis. In the parametrization, to reduce computational time, radiative transfer simulations are run for a single view factor weighted average path length representing the average distance from the surface to the sensor as "seen" by the sensor over its FOV. This view factor weighted "hemispherical" path length can be calculated using a sensor view model (such as SUM) or approximated by hand using a digital building model (although care should be taken to ensure the sensor cosine response is accurately represented). Using the view factor weighted path length, a suite of radiative transfer simulations is run prior to correction to retrieve "hemispherical" broadband atmospheric transmittances for a range of potential humidities. Humidity-transmittance pairings are aggregated into a look-up table (LUT), where the measured humidity at each time step can be matched with the associated hemispherical transmittance. Using the hemispherical transmittance as the fraction of surface emitted L\textsubscript{up} reaching the sensor, measured L\textsubscript{up} can then be decomposed into its surface and atmospheric emitted components, from which a parameterized T\textsubscript{hem, r} (T\textsubscript{hem, para}) is inferred. Comparison of T\textsubscript{hem, r} derived from the correction method and the parameterization show good agreement, the results of which are included in Figure \ref{para_eval} and Table \ref{parastats}.

@@

\begin{table}[H]
	\centering
	\caption{A description of morphological parameters and measured variables for BUBBLE urban and rural sites. Modified from \citet{Rotach2005} to include only relevant parameters.}
	\label{morphbspr}
	\begin{tabular*}{\textwidth}{p{3.75cm} p{2.25cm}p{3.5cm}p{2.75cm}p{2.75cm}}
		\toprule 
		Site & Location & Morphological & Meteorological & Radiation \\ 
		& Height & Characteristics\footnote{} & Variables\footnote{} & [No of levels]\footnote{} \\ 	\midrule

@@

sUHI and clUHI magnitudes were calculated from T\textsubscript{hem, r} and T\textsubscript{air} (measured from 2 \si{\meter} above ground level) as $\Delta$T\textsubscript{(urban-rural)} at 30 \si{\min} intervals. Rural T\textsubscript{hem, b} was calculated an expanded version of equation \ref{stefb1} to remove emissivity effects,

@@

\noindent where $L_z^\uparrow$ is at-sensor upwelling longwave radiation, $L_{sky}^\downarrow$ is at-sensor reflected downwelling longwave radiation, and emissivity ($\epsilon$) = 0.94. Sensitivity tests included in Section \ref{sensitivity} of the companion paper show that screen level irradiances ($ z $ = ~2 \si{\meter}) are not subject to significant effects from the intervening atmosphere and do not require atmospheric correction. Rural irradiances in this study were measured from approximately 2 \si{\meter} above ground, thus rural T\textsubscript{hem, b} is approximately equal to T\textsubscript{hem, r}.

@@

Figure \ref{meteo_wv_sol} panel (a) shows mean and maximum sUHI magnitudes as a function of solar radiant exposure over each day of the summer months of the BUBBLE campaign. Point shading indicates the month (white is April and black is July). A clear positive relationship is observed between accumulated solar radiation and both sUHI\textsubscript{max} and sUHI\textsubscript{mean}. Panel (b) of Figure \ref{meteo_wv_sol} shows mean and maximum hemispherical sUHI magnitudes as a function of mean wind velocity for the duration of the eight month BUBBLE climatology. Colors indicate solar radiant exposure accumulated over each day. Similar results are observed when rural winds are substituted.

Figures \ref{meteo_hum} and \ref{meteo_atuhi} show hemispherical sUHI magnitudes as a function of water vapor content and clUHI measured over the BUBBLE campaign. Winter hours are omitted in both as sUHI is small and displays minimal variation in winter. Jamie: still working on getting color palettes for these. Will merge into one figure.

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_wv_sol}
	\caption{(a) Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day of April through July versus integrated solar radiant exposure measured over the same 24 \si{\hour} period. Shading indicates month of the year, with dark points indicating later months. (b) Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day in the eight month climatology versus mean wind velocity measured at approximately 2 \si{\meter} above ground at the Sperrstrasse site. Coloring indicates integrated solar radiant exposure over each day.}
	\label{meteo_wv_sol}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_hum}
	\caption{sUHI magnitude versus atmospheric water vapor content measured at 2 \si{\meter} at the Sperrstrasse site. Similar patterns are observed when rural water vapor content is substituted.}
	\label{meteo_hum}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_atuhi}
	\caption{sUHI magnitude versus clUHI magnitude. clUHI is calculated as the difference between urban and rural T\textsubscript{air} measured at 2 \si{\meter} above ground level.}
	\label{meteo_atuhi}
\end{figure}

@@

\subsection{The effect of rural/non-urban characteristics on sUHI}

@@

@@

\section{Discussion}
\citet{Oke2017} presents a generalized conceptual model for sUHI development for multiple representations of the surface. However, the results in \citet{Oke2017} are based on simulations of an idealized urban area without trees, complex non-orthogonal surface geometry, or sloped facets and have not been supported observationally. The results in this chapter provide an ideal platform to test conceptual understanding of sUHI development, to understand the diurnal and seasonal nature of sUHI and the meteorological factors that drive its development.

@@

sUHI development is the result of urban modification to surface geometry and thermal and radiative properties, most notably sky-view-factor, impervious surface fraction, and thermal admittance (and by relation, urban-rural soil moisture contrasts). These differences alter urban heating/cooling rates relative to non-urban spaces and provide the genesis for diurnal sUHI development. In the early morning hours, rural surfaces respond quickly to insolation at low solar angles, and rural T\textsubscript{surf} rapidly increases. In contrast, shading by urban features causes urban T\textsubscript{surf} to respond more slowly, resulting in a small sUHI\textsubscript{min} in the hours just after sunrise. This is observed in Basel across all representations of urban T\textsubscript{surf}. Through the late morning hours, urban heating rates increase relative to rural rates as the solar zenith angle decreases and low thermal admittance facets (particularly rooftops) experience more direct insolation and quickly heat up. Increases in urban T\textsubscript{surf} are amplified by the abundance of impervious surfaces, which result in a smaller fraction of insolation partitioned to latent heat relative to vegetated non urban spaces. After the initial spike in rural heating in the early morning hours, urban heating rates overtake rural approximately three to four hours after sunrise and the sUHI slowly grows through midday to a peak at approximately two to three hours hours after solar noon. In the hours after sUHI\textsubscript{max}, the solar zenith angle increases and both urban and rural heating rates decline and begin to converge, as shading by canyon geometry causes some facets to begin to cool.

In the afternoon hours, both urban and rural surfaces transition to cooling; this can lead to several different patterns in evening sUHI development and is strongly controlled by urban surface structure. Relative to non-urban spaces, cooling rates for wall and road facets lag slightly as a result of canyon radiation trapping, this is offset by increased cooling rates for low thermal admittance rooftop facets. These two factors control whether sUHI grows or shrinks in the hours after sUHI\textsubscript{max}. In urban areas with relatively unobstructed sky-view-factors or large rooftop plan area fractions, strong radiative cooling can lead to a sharp reduction in sUHI. The opposite is true of urban areas with restricted sky view factors and tall street canyons with large wall surface area fractions, where canyon radiative trapping reduces radiative cooling and sUHI is maintained through the evening. For the Sperrstrasse canyon, the combination of a narrow but relatively low street canyon results in a slight reduction in Basel's sUHI as strong radiative cooling of rooftops is slightly offset by canyon radiation trapping. The dependence of sUHI development on canyon geometry is illustrated well by comparing nadir and complete representations of sUHI observed in Basel. A nadir view, by virtue of oversampling rapidly cooling rooftops, observes a strong evening decline in sUHI. In contrast, by sampling vertical facets that cool more gradually, this decline is less evident in sUHI calculated from a complete view of Basel T\textsubscript{surf}. Thus, as expected, remote sensed sUHI analysis is also directionally dependent and subject to anisotropic effects when viewed from a narrow-FOV sensor. These effects not only affect the magnitude of remote sensed sUHI, but may also alter perceived diurnal patterns of sUHI.

After sunset, the unobstructed rural sky-view-factor allows for strong radiative cooling of the surface and rural cooling rates spike relative to urban, leading to a slight jump in sUHI. However, variations in rural thermal admittance, surface geometry, land cover and use, and/or influence from surrounding terrain can result in significant contrasts in rural cooling rates. All of these factors show strong spatiotemporal variation, resulting in large differences in sUHI development depending on rural site characteristics, similar to observations of variations in clUHI based on rural siting in \citep{Runnalls2000}. This is illustrated by two different patterns of post-sunset sUHI development observed for the two rural sites in Basel. Although similar in terms of surface coverage and site geometry, hilly terrain surrounding the Lange Erlen site results in suspected cold air drainage and an increased rural cooling rate, particularly following hot, clear sky days. This effect is not present at the Village Neuf site, where nocturnal cooling rates are likely driven primarily driven by radiative losses. The "extra", largely non-radiative cooling at the Lange Erlen site results in a strong second peak in sUHI that is not present in sUHI calculated using the Village Neuf site. Following site dependent effects on sUHI in the hours after sunset, sUHI in Basel is relatively constant over the nighttime hours, with slight growth observed on clear nights where canyon radiation trapping retards urban nocturnal cooling rates relative to rural.

@@

sUHI is highly dependent on meteorological conditions, which account for much of the day to day variation sUHI magnitudes. Conditions that foster the largest microscale spatiotemporal contrasts in urban T\textsubscript{surf} also entail the largest sUHI magnitudes. Contra, conditions that suppress microscale contrasts in T\textsubscript{surf} show reduced sUHI magnitudes. As such, sUHI development is strongly tied to integrated solar radiant exposure accumulated over a day and wind velocity, which display positive and negative relationships with sUHI magnitudes respectively. Meteorological controls on sUHI are summarized in Table \ref{meteo_cont}.

@@

As was found with correction magnitudes in Chapter \ref{paper1}, sUHI magnitudes are not strongly correlated with water vapor content as, in the absence of fog, humidity near the surface does not strongly influence surface heating or cooling rates in and of itself. Atmospheric transmittance of longwave radiation decreases sharply with small increases in water vapor when the atmosphere is very dry - shown in Figure \ref{humtest}. As water vapor content increases from zero, variation in spectral atmospheric transmittance decreases. Therefore, typical variations in urban and rural humidities have little effect on the fraction of outgoing longwave radiation available for heating the canopy layer air volume through absorption and for heating of the surface by atmospheric emission. As a result, water vapor content has little effect on sUHI.

\begin{table}[H]
	\centering
	\caption{A summary of the relationship between sUHI, meteorological variables, and clUHI.}
	\label{meteo_cont}
	\begin{tabular}{lc}
		\toprule 
		An increase in $x$, has a/an... & $y$ effect on sUHI magnitude. \\
		\midrule
		$\Delta$T\textsubscript{surf - air}, \si{\kelvin} &\textit{increasing} \\
		&\\
		T\textsubscript{surf}, \si{\degreeCelsius}  & \textit{increasing} \\
		&\\
		T\textsubscript{air}, \si{\degreeCelsius} & \textit{increasing} \\
		&\\
		Insolation, \si{\mega\joule\per\meter\squared\per\day} & \textit{increasing}  \\
		& \\
		Water vapor content, \si{\gram\per\meter\squared} & \textit{negligible} \\
		&\\
		Wind velocity, \si{\meter\per\second} &\textit{decreasing} \\
		&\\
		clUHI, \si{\kelvin} &\textit{increasing} \\
		\bottomrule
	\end{tabular} 
\end{table}

>>>>

\subsection{Method evaluation and sensitivity testing}

@@

Sensor placement sensitivity testing in section \ref{sensitivity} shows that T\textsubscript{hem, r} is spatially variant when measured over urban areas at typical pyrgeometer heights. This introduces a potential bias in T\textsubscript{hem, r} as changing sensor position relative to canyon features alters view factor proportions of rooftops, roads, and walls. The sensitivity of T\textsubscript{hem, r} to sensor placement decreases with height and T\textsubscript{hem, r} becomes spatially invariant at approximately 3.5 times mean building height \citep{Roberts2010,Adderley2015}. As such, these biases can be significantly reduced by mounting downward facing pyrgeometers slightly higher than normal. Although pyrgeometers at both the Sperrstrasse and are mounted below the height at which T\textsubscript{hem,r} is spatially invariant, sensitivity tests in section \ref{sensitivity} show that the magnitude of this bias is relatively small for sensor placement at the Sperrstrasse canyon. Similar or lesser biases as the sensor is mounted offset from surrounding rooftops (which contribute to large diurnal amplitude in T\textsubscript{hem, r} and larger deviation from T\textsubscript{comp}). In addition, T\textsubscript{hem, r} shows good performance relative to T\textsubscript{plan} when compared to T\textsubscript{comp}, particularly under clear sky conditions when microscale spatiotemporal variations in urban T\textsubscript{surf} and sUHI magnitudes are greatest. As such, the influence of biases in T\textsubscript{hem} in sUHI analysis for both BUBBLE and EPiCC campaigns is likely small.

\subsection{A parameterization scheme}

By virtue of requiring nearly 500 radiative transfer simulations per time step and significant post-processing, the method described in \ref{method2} is computationally expensive. This prompted the development of a parameterization scheme to facilitate more widespread usage of urban T\textsubscript{hem, r} for climatological and comparative sUHI analysis. In the parametrization, to reduce computational time, radiative transfer simulations are run for a single view factor weighted average path length representing the average distance from the surface to the sensor as "seen" by the sensor over its FOV. This view factor weighted "hemispherical" path length can be calculated using a sensor view model (such as SUM) or approximated by hand using a digital building model (although care should be taken to ensure the sensor cosine response is accurately represented). Using the view factor weighted path length, a suite of radiative transfer simulations is run prior to correction to retrieve "hemispherical" broadband atmospheric transmittances for a range of potential humidities. Humidity-transmittance pairings are aggregated into a look-up table (LUT), where the measured humidity at each time step can be matched with the associated hemispherical transmittance. Using the hemispherical transmittance as the fraction of surface emitted L\textsubscript{up} reaching the sensor, measured L\textsubscript{up} can then be decomposed into its surface and atmospheric emitted components, from which a parameterized T\textsubscript{hem, r} is inferred. Comparison of T\textsubscript{hem, r} derived from the correction method and the parameterization show good agreement, the results of which are included in Figure \ref{para_eval} and Table \ref{parastats}.

@@

\begin{table}[H]
	\centering
	\caption{A description of morphological parameters and measured variables for BUBBLE urban and rural sites. Modified from Rotach et al. (2005) to include only relevant parameters.}
	\label{morphbspr}
	\begin{tabular*}{\textwidth}{p{3.75cm} p{2.25cm}p{3.5cm}p{2.75cm}p{2.75cm}}
		\toprule 
		Site & Location & Morphological & Meteorological & Radiation \\ 
		& Height & Characteristics\footnote{} & Variables\footnote{} & [No of levels]\footnote{} \\ 	\midrule

@@

sUHI and clhUHI magnitudes were calculated from T\textsubscript{hem, r} and T\textsubscript{air} (measured from 2 \si{\meter} above ground level) as $\Delta$T\textsubscript{(urban-rural)} at 30 \si{\min} intervals. Rural T\textsubscript{hem, b} was calculated an expanded version of equation \ref{stefb1} to remove emissivity effects,

@@

\noindent where $L_z^\uparrow$ is at-sensor upwelling longwave radiation, $L_{sky}^\downarrow$ is at-sensor reflected downwelling longwave radiation, and rural emissivity ($\epsilon$) = 0.94. Sensitivity tests included in Section \ref{sensitivity} of the companion paper show that screen level irradiances ($ z $ = ~2 \si{\meter}) are not subject to significant effects from the intervening atmosphere and do not require atmospheric correction. Rural irradiances in this study were measured from approximately 2 \si{\meter} above ground, thus rural T\textsubscript{hem, b} is approximately equal to T\textsubscript{hem, r}.

@@

Figure \ref{meteo_sol} shows mean and maximum sUHI magnitudes (sUHI\textsubscript{mean} and sUHI\textsubscript{max}) as a function of solar radiant exposure for each day of the BUBBLE campaign. Point shading indicates the season, where black is winter (Nov - Feb), white is spring (Mar - Apr), and red is summer (May - Aug). A clear positive relationship is observed between accumulated solar radiation and both sUHI\textsubscript{max} and sUHI\textsubscript{mean}.

Figure \ref{meteo_wv} shows hemispherical sUHI\textsubscript{mean} and sUHI\textsubscript{max} magnitudes as a function of mean wind velocity for each day of the BUBBLE climatology. A weak negative relationship is observed between mean wind velocity and sUHI\textsubscript{max}, where larger sUHI\textsubscript{max} occurs under low wind conditions. sUHI\textsubscript{mean} yields a slightly more muted relationship. Similar results are observed when rural winds are substituted. Figure \ref{meteo_hum} shows hemispherical sUHI\textsubscript{mean} and sUHI\textsubscript{max} magnitudes as a function of mean atmospheric water vapor content for each day over the BUBBLE campaign. No discernible relationship is observed between atmospheric water vapor content and sUHI\textsubscript{mean} or sUHI\textsubscript{max}.  In both, colors indicate solar radiant exposure accumulated over each day (yellow is large accumulated insolation, brown is low).

Figured \ref{meteo_atuhi} shows hemispherical sUHI magnitudes as a function of water vapor content and clUHI measured over the summer months of the BUBBLE campaign (May through July 2002). Winter hours are omitted as sUHI is small and displays minimal variation in winter. The relationship between sUHI and clUHI is complex and highly dependent on time of day - owing largely to similar, but slightly offset patterns of sUHI and clUHI development. Coloring in the daytime plot indicates time of day: Red is morning (5:00 - 9:00 LST), gray is midday hours (9:30 - 17:00 LST), and yellow is evening hours (17:30 - 20:00 LST).

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_kdown}
	\caption{Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day of April through July versus integrated solar radiant exposure measured over the same 24 \si{\hour} period. Color indicates season: black is winter (Nov - Feb), white is spring (Mar - Apr), and red is summer (May - Aug). }
	\label{meteo_sol}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_wv}
	\caption{Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day versus mean wind velocity measured at approximately 2 \si{\meter} above ground at the Sperrstrasse site. Coloring indicates integrated solar radiant exposure over each day.}
	\label{meteo_wv}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_hum}
	\caption{Maximum and mean 24 \si{\hour} sUHI magnitude versus atmospheric water vapor content measured at 2 \si{\meter} at the Sperrstrasse site. Similar patterns are observed when rural water vapor content is substituted. Coloring indicates integrated solar radiant exposure over each day.}
	\label{meteo_hum}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{meteo_atuhi}
	\caption{sUHI magnitude versus clUHI magnitude binned for day and night calculated over the summer months of the BUBBLE campaign.  Coloring in the daytime plot indicates time of day: Red is morning (5:00 - 9:00 LST), gray is midday hours (9:30 - 17:00 LST), and yellow is evening hours (17:30 - 20:00 LST).}
	\label{meteo_atuhi}
\end{figure}

%Vancouver

Figure \ref{vanc_sol} shows mean and maximum sUHI magnitudes (sUHI\textsubscript{mean} and sUHI\textsubscript{max}) as a function of solar radiant exposure over each day of the year long EPiCC Vancouver campaign. Point shading indicates the season, where black is winter (Nov - Feb), white is spring (Mar - Apr), red is summer (May - Aug), and blue is fall (Sept - Oct). A clear positive relationship is observed between accumulated solar radiation and both sUHI\textsubscript{max} and sUHI\textsubscript{mean}, with significant seasonal variation.

Figure \ref{vanc_wv} shows hemispherical sUHI\textsubscript{mean} and sUHI\textsubscript{max} magnitudes as a function of mean wind velocity for each day of the EPiCC Vancouver campaign. A weak negative relationship is observed between mean wind velocity and both sUHI\textsubscript{mean} and sUHI\textsubscript{max}, where larger sUHI magnitudes tend to occur under low wind conditions. Higher overall sUHI magnitudes and mean wind velocities in Vancouver make the relationship between sUHI and mean wind velocity slightly more evident. Similar results are observed when rural winds are substituted. Figure \ref{vanc_hum} shows hemispherical sUHI\textsubscript{mean} and sUHI\textsubscript{max} magnitudes as a function of mean atmospheric water vapor content for each day of the EPiCC Vancouver campaign. In contrast to findings in the BUBBLE campaign, a weak relationship between sUHI and humidity is observed. However, the apparent relationship may be a product of Vancouver's proximity to the coast, as there is significant seasonal variation in both  and may be no actual causal relationship. In both, colors indicate solar radiant exposure accumulated over each day (yellow is large accumulated insolation, brown is low).

Figured \ref{vanc_swc} shows hemispherical sUHI magnitudes as a function of water vapor content and clUHI measured for April and May 2009 of the EPiCC Vancouver canpaign. These months were selected to show a wide range of soil water contents and sUHI magnitudes. A negative relationship is observed between sUHI and soil water vapor content, particularly in May, where larger sUHI magnitudes often coincide with lower soil water vapor contents. However, some of this relationship may owe to diurnal patterns of soil saturation, which, contrary to sUHI magnitudes, tend to be highest in early morning and lower throughout the day. Color in both plots indicates time of day: shades of blue indicate early morning through midday hours, white is solar noon through early evening hours, and red is late evening through the nighttime.

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{vanc_kdown}
	\caption{Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day of the EPiCC climatology versus integrated solar radiant exposure measured over the same 24 \si{\hour} period. Color indicates season: black is winter (Nov - Feb), white is spring (Mar - Apr), red is summer (May - Aug), and blue is fall (Sept - Oct).}
	\label{vanc_sol}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{vanc_wv}
	\caption{Maximum and mean 24 \si{\hour} hemispherical sUHI magnitude for each day over the EPiCC climatology versus mean wind velocity measured at approximately 10 \si{\meter} above ground at the Sunset tower site. Similar patterns are observed when rural mean wind velocity is substituted. Coloring indicates integrated solar radiant exposure over each day.}
	\label{vanc_wv}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{vanc_hum}
	\caption{Maximum and mean 24 \si{\hour} sUHI\textsubscript{hemi} magnitude versus mean atmospheric water vapor content measured at 2 \si{\meter} at the Sunset tower urban site in Vancouver, Canada. Similar patterns are observed when rural water vapor content is substituted. Coloring indicates integrated solar radiant exposure over each day.}
	\label{vanc_hum}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{vanc_swc}
	\caption{sUHI\textsubscript{hemi} magnitude versus soil water content measured from 0.05 \si{\meter} below the surface at the Westham island site in Vancouver, Canada for April and May 2009. Color indicates time of day: blue is early morning through midday, white is solar noon through early evening, and red is late evening through the night.}
	\label{vanc_swc}
\end{figure}

\subsection{The effect of rural/non-urban characteristics on sUHI and a generalizable case for sUHI}

@@

Figure \ref{gen_case} presents a schematic for summertime sUHI development constructed from mean urban and rural T\textsubscript{hem} over the summer months of the EPiCC Vancouver campaign. sUHI shows clear diurnal dependence on differential heating/cooling rates at the urban and rural sites, with greater urban heating rates observed from sunrise to approximately solar noon forcing the development of sUHI and greater urban cooling rates in the afternoon and evening hours forcing the destruction of daytime sUHI. These results are highly generalizable to other mid-latitude cities.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=15cm,height=7in,keepaspectratio]{gen_case}
	\caption{Top: Mean T\textsubscript{hem, r} for the Sunset tower (urban) and Westham island (rural) sites in Vancouver, Canada over May - Aug 2009. Middle: Mean heading/cooling rates for urban and rural sites. Bottom: Mean sUHI\textsubscript{hemi} and hourly rate of change in sUHI\textsubscript{hemi}.  n = 5905}
	\label{gen_case}
\end{figure}

\section{Discussion}
\citet{Oke2017} presents a generalized conceptual model for sUHI development for multiple representations of the surface. However, the results in \citet{Oke2017} are based on simulations of an idealized urban area without trees, complex non-orthogonal surface geometry, or sloped facets and have not been supported observationally. The results in this chapter provide an ideal platform to test conceptual understanding of sUHI development, to understand and characterize the diurnal and seasonal nature of sUHI, and to elucidate meteorological and climatological factors that drive its development.

@@

sUHI development is the result of urban modification to surface geometry and thermal and radiative properties, most notably sky-view-factor, impervious surface fraction, and thermal admittance (and by relation, urban-rural soil moisture contrasts). These differences alter urban heating/cooling rates relative to non-urban spaces and provide the genesis for diurnal sUHI development. In the early morning hours, rural surfaces respond quickly to insolation at low solar angles, and rural T\textsubscript{surf} rapidly increases. In contrast, shading by urban features causes urban T\textsubscript{surf} to respond more slowly, resulting in a small sUHI\textsubscript{min} in the hours just after sunrise. This is observed in Basel across all representations of urban T\textsubscript{surf} and in the generalizable case presented for Vancouver sUHI. Through the late morning hours, urban heating rates increase relative to rural rates as the solar zenith angle decreases and low thermal admittance facets (particularly rooftops) experience more direct insolation and quickly heat up. Increases in urban T\textsubscript{surf} are amplified by the abundance of impervious surfaces, which result in a smaller fraction of insolation partitioned to latent heat relative to vegetated non urban spaces. After the initial spike in rural heating in the early morning hours, urban heating rates overtake rural approximately three to four hours after sunrise and the sUHI slowly grows through midday to a peak at approximately two hours hours after solar noon. In the hours after sUHI\textsubscript{max}, the solar zenith angle increases and both urban and rural heating rates decline and begin to converge, as shading by canyon geometry causes some facets to begin to cool.

In the afternoon hours, both urban and rural surfaces transition to cooling; this can lead to several different patterns in evening sUHI development and is strongly controlled by urban surface structure. Relative to non-urban spaces, cooling rates for wall and road facets lag slightly as a result of canyon radiation trapping, this is offset by increased cooling rates for low thermal admittance rooftop facets. These two factors control whether sUHI grows or shrinks in the hours after sUHI\textsubscript{max}. In urban areas with relatively unobstructed sky-view-factors or large rooftop plan area fractions, strong radiative cooling can lead to a sharp reduction in sUHI. The opposite is true of urban areas with restricted sky view factors and tall street canyons with large wall surface area fractions, where canyon radiative trapping reduces radiative cooling and sUHI is maintained through the evening. For the Sperrstrasse canyon, the combination of a narrow but relatively low street canyon results in a slight reduction in Basel's sUHI as strong radiative cooling of rooftops is slightly offset by canyon radiation trapping. In Vancouver, urban cooling rates quickly overtake rural - owing to strong radiative cooling - and sUHI declines precipitously after sUHI\textsubscript{max}. The dependence of sUHI development on canyon geometry is illustrated well by comparing nadir and complete representations of sUHI observed in Basel. A nadir view, by virtue of oversampling rapidly cooling rooftops, observes a strong evening decline in sUHI. In contrast, by sampling vertical facets that cool more gradually, this decline is less evident in sUHI calculated from a complete view of Basel T\textsubscript{surf}. Thus, as expected, remote sensed sUHI analysis is also directionally dependent and subject to anisotropic effects when viewed from a narrow-FOV sensor. These effects not only affect the magnitude of remote sensed sUHI, but may also alter perceived diurnal patterns of sUHI.

After sunset, the unobstructed rural sky-view-factor allows for strong radiative cooling of the surface and rural cooling rates spike relative to urban, leading to a slight jump in sUHI. This is observed in Vancouver sUHI as a slight reduction in the rate of sUHI decline near 20:00 LST. However, variations in rural thermal admittance, surface geometry, land cover and use, and/or influence from surrounding terrain can result in significant contrasts in rural cooling rates. All of these factors show strong spatiotemporal variation, resulting in large differences in sUHI development depending on rural site characteristics, similar to observations of variation in clUHI based on rural siting in \citep{Runnalls2000}. This is illustrated by two different patterns of post-sunset sUHI development observed for the two rural sites in Basel. Although similar in terms of surface coverage and site geometry, hilly terrain surrounding the Lange Erlen site results in suspected cold air drainage and an increased rural cooling rate, particularly following hot, clear sky days. This effect is not present at the Village Neuf site, where nocturnal cooling rates are likely driven primarily driven by radiative losses. The "extra", largely non-radiative cooling at the Lange Erlen site results in a strong second peak in sUHI that is not present in sUHI calculated using the Village Neuf site. In the context of the Lowry method discussed in Chapter \ref{intro}, $V_{M, \ R}$ measured from the Lange Erlen site and $V_{M, \ U}$ are not subject to the same $V_{L}$ and equation does not yield a perfectly isolated $V_{H}$ in the hours just after sunset.

Following site dependent effects on sUHI in the hours after sunset, sUHI in Basel is relatively constant over the nighttime hours, with slight growth observed on clear nights where canyon radiation trapping retards urban nocturnal cooling rates relative to rural. In contrast, mean sUHI in Vancouver decreases through the night as the urban area experiences slight cooling. Both of these cases demonstrate the dependence of observed sUHI on urban and rural site characteristics and sensor-surface geometry illustrating the need for proper critical analysis of the ensemble of climactic forcings present at a given urban-rural pairing and the potential for influence from sensor or instrument bias.

@@

sUHI is highly dependent on meteorological conditions, which account for much of the day to day variation sUHI magnitudes. Conditions that foster the largest microscale spatiotemporal contrasts in urban T\textsubscript{surf} also entail the largest sUHI magnitudes. Contra, conditions that suppress microscale contrasts in T\textsubscript{surf} show reduced sUHI magnitudes. As such, sUHI development is strongly tied to integrated solar radiant exposure accumulated over a day and wind velocity, which display positive and negative relationships with sUHI magnitudes respectively. Other variables also show weakly positive (mean atmospheric water vapor content) and negative (soil water content) relationships with sUHI, however, conclusions about causality are more difficult to make as they too display some dependence on the same variables. Meteorological controls on sUHI are summarized in Table \ref{meteo_cont}.

@@

sUHI magnitudes in the BUBBLE and EPiCC campaigns yield slightly different relationships with atmospheric water vapor content, with no relationship observed in Basel and a weakly positive one observed in Vancouver. It is unlikely that these are a result of any significant physical or causal connection between the two variables as, in the absence of fog, humidity near the surface does not strongly influence surface heating or cooling rates in and of itself. Atmospheric transmittance of longwave radiation decreases sharply with small increases in water vapor when the atmosphere is very dry, with only marginal subsequent decreases - shown in Figure \ref{humtest}. As water vapor content increases from zero, variation in spectral atmospheric transmittance decreases. Therefore, typical variations in urban and rural humidities have little effect on the fraction of outgoing longwave radiation available for heating the canopy layer air volume through absorption and for heating of the surface by atmospheric emission. As a result, a similar relationship is expected to that found with correction magnitudes in Chapter \ref{paper1} is expected between humidity and sUHI. A weak positive correlation between the two is therefore likely a result of coincident seasonality in both variables as both mean daytime sUHI and mean water vapor mass densities are large in summer and small in winter. As this seasonality is evident in both Basel and Vancouver, it is likely that a full climatology of sUHI and humidity would yield similar correlation (but not causation) in Basel as is observed in Vancouver.

\begin{table}[H]
	\centering
	\caption{A summary of the relationship between sUHI, meteorological variables, and clUHI.}
	\label{meteo_cont}
	\begin{tabular}{lc}
		\toprule 
		An increase in $x$, has a/an... & $y$ effect on sUHI magnitude. \\
		\midrule
		$\Delta$T\textsubscript{surf - air}, \si{\kelvin} &\textit{increasing} \\
		&\\
		T\textsubscript{surf}, \si{\degreeCelsius}  & \textit{increasing} \\
		&\\
		T\textsubscript{air}, \si{\degreeCelsius} & \textit{increasing} \\
		&\\
		Insolation, \si{\mega\joule\per\meter\squared\per\day} & \textit{increasing}  \\
		& \\
		Atmospheric water vapor content, \si{\gram\per\meter\cubed} & \textit{negligible} \\
		&\\
		Volumetric soil moisture content, \si{\percent} & \textit{decreasing} \\
		&\\
		Wind velocity, \si{\meter\per\second} &\textit{decreasing} \\
		&\\
		clUHI, \si{\kelvin} &\textit{increasing} \\
		\bottomrule
	\end{tabular} 
\end{table}
